The effects of binarity on planet occurrence rates measured by transit surveys

### 2018-02-14 paper/main.tex
improve text
substantially

Wide-field surveys for transiting planets, such as the NASA
{\it Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars are binaries. Unresolved and unrecognized binary
stars give rise to systematic errors in planet occurrence rates,
including misclassified planets and miscounts in the number of
searched stars. The individual errors can have different signs,
making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive
a formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula
depends on the binary fraction; the mass function of the secondary
stars; and the true occurrence of planets around primaries,
secondaries, and single stars. It also takes into account the
Malmquist bias by which binaries are over-represented in
flux-limited samples. Application of the formula to an idealized
{\it Kepler}-like survey shows that for planets larger than
$2r_\oplus$, the net systematic error is of order 10\%. For smaller
planets the errors are potentially larger: the occurrence of
Earth-sized planets could be overestimated by as much as a factor of
two. One consequence is that unrecognized binaries cannot account
for the apparent discrepancy between hot Jupiter occurrence
rates measured by transit and RV surveys. We also show that if
high-resolution imaging reveals a transit host star to be a binary,
the planet is usually more likely to orbit the primary star than the
secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern in occurrence rate studies is that in most cases,
investigators have assumed that all of the sources of light that were
monitored are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. In reality many of
them are unresolved multiple-star systems, especially binaries.
Unrecognized binaries cause numerous systematic errors in the
planetary occurrence rates.  For example, when there is a transiting
planet around a star in a binary, the additive constant light from the
second star reduces the fractional loss of light due to the planet.
This ``flux dilution'' makes transit signals harder to detect and
lowers the number of detections.  On the other hand, in a binary there
two opportunities to detect transiting planets, which could increase
the overall number of detections.

At the outset of this study it was not clear to us whether the neglect
of binaries was a serious issue at all, or even whether the net effect
of all the errors is positive or negative.  The goal of this study was
to clarify the various sources of error and provide a framework for
dealing with these issues.  In this spirit, most of our models are
idealized and analytic, and we do not attempt a detailed correction of
the results from {\it Kepler} or any other real transit survey.
Our goal is to estimate the order of magnitude of the effects,
assuming the survey is signal-to-noise limited, and given the true
underlying planet occurrence rates.

@@

Imagine a group of astronomers that wants to measure the mean number of
planets of a certain type per star of a certain type.  They perform a
wide-field photometric search for planets that transit, and discover
all for which
\begin{equation}
\frac{\delta_{\rm obs}}{\sigma}
  = ({\rm constant}) \cdot \delta_{\rm obs} L_{\rm sys}^{1/2} d^{-1}
> \left(\frac{{\rm S}}{{\rm N}}\right)_{\rm min}.
\label{eq:S_N_thresh}
\end{equation}
Here the signal S is the observed transit depth $\delta_{\rm obs}$, in
units of relative flux.
%LB: we don't introduce r or R because they are system-specific.
%For single stars, $\delta_{\rm obs} \approx (r/R)^2$; for binaries,
%flux dilution $\mathcal{D}$ lowers the observed depth to
%$\delta_{\rm obs} \approx \mathcal{D} (r/R)^2$. 
The noise N is the uncertainty in the determination of the flux ratio
of the star inside and outside of transits.  Throughout this paper we
will assume the photometric uncertainty to scale as the inverse square
root of the number of photons collected from the source, {\it i.e.},
the dependence on stellar properties is $\sigma \propto (L_{\rm
sys}/d^2)^{-1/2}$, where $L_{\rm sys}$ is the luminosity of the
stellar system and $d$ is its distance from Earth.
% The solid angle coverage, telescope area, and
% survey duration are in the ``constant'' term of
% Equation~\ref{eq:S_N_thresh}.  The transit duration also affects
% detectability, but we omit it in this work for brevity.

The astronomers then assume that all stars are single.  In particular
they do not have accurate enough parallaxes to tell that some of the
stars are apparently overluminous.  They want to compute the number of
planets with size $r$ per star of mass $M$.  If the total number of
stars for which such planets could be detected ($N_\star$) and the
number of detected planets ($N_{\rm det}$) are both large, then the
astronomers' estimate for the occurrence rate ($\Lambda$) is simply
\begin{equation}
\Lambda = \frac{N_{\rm det}}{N_\star}
                    \times \frac{1}{p_{\rm tra}},
\label{eq:occ_rate_simple}
\end{equation}
where $p_{\rm tra}$ is the geometric transit probability, needed to
account for the fact that most planetary orbits are not aligned close
enough with our line of sight to produce transits.

@@

We will focus exclusively on the subset of problems that arise from
the fact that many stars exist in gravitationally bound binary
systems.  Even with this narrow focus, there are numerous sources of
error.  All three of the quantities in
Equation~\ref{eq:occ_rate_simple} are subject to observational bias:
\begin{enumerate}
%    
    \item The number of planets, $N_{\det}$, is actually the number of
      detected planets that {\it appear} to have size $r$, orbiting
      stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size will be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The apparent number of stars that were searched, $N_\star$,
      is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were searched for transiting
          planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be suitably small and bright to detect the
          desired type of planet are in fact binaries for which the
          amplitude of the photometric signal would be reduced to an
          undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$, which scales with the
      stellar mean density as $\rho^{-1/3}$, is biased because the
      planet-hosting star could be misclassified.
%    
\end{enumerate}

Two other problems that may arise are not represented in
Equation~\ref{eq:occ_rate_simple}.  One is astrophysical: the true
occurrence rate of the desired type of planet may depend on whether
the host is a single star, the primary star of a binary, or the
secondary star of a binary.  Such differences could be caused by the
requirement for long-term dynamical stability, or differences in the
planet formation process.  When the search sample includes both
singles and binaries, the detected planets are thereby drawn from
different occurrence 
distributions~\citep[{\it e.g.},][]{
    wang_occurrence_2015,kraus_impact_2016}.

The other problem is an observational effect.  Even after the
astronomers learn about binaries and attempt to correct for their
presence, they must realize that the ratio of binaries to single stars
in the searched sample will differ from that in a volume-limited
sample due to a type of Malmquist bias.  Binaries have a higher total
luminosity than either the primary or secondary star would have on its
own.  This means that the binaries are searchable for transit signals
of a given amplitude at greater distances from the Earth.  The
binaries in a sample of apparently searchable stars are therefore
drawn from a larger search volume than the single stars.

@@

Given all of the confusing and opposing sources of error, we will
proceed in stages. We start from a barebones model in which everything
can be written down on the back of a napkin, and build up to an
analytic model allowing for generality in the distribution of the
binaries and the planets they host.  We consider only corrections to
the occurrence as a function of planetary radius, and not orbital
period (the other key variable in transit studies), because when
multiple transits are detected the period can be directly measured
regardless of whether the host star is single or part of a binary.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{
        Searchable volumes for single stars ({\it left}) and twin
        binaries ({\it right}), assuming that all stars have the same
        mass, radius, and luminosity, and that all planets have the
        same radius $r=r\p$.
        {\it Top:} At an apparent radius $r\a=r\p$, the observer
        searched twin binaries out to a distance $\sqrt{2}\times$ that
        of single stars.  Because of dilution, there are no planets in
        twin binaries with $r\a=r\p$.
        {\it Bottom:} At $r\a=r\p/\sqrt{2}$, the maximum searchable
        distances are half those at $r\a=r\p$.  The only detected
        planets with $r\a=r\p/\sqrt{2}$ orbit twin binaries.
    }
    \label{fig:model_1_volumes}
\end{figure}

Since the effects of binarity are most pronounced when the two
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Some fraction of them exist in twin binaries. All planets have
size $r\p$.  In this scenario, stars are never misclassified because
the combined light of a binary has the same color and spectrum as a
single star.  We further assume that planets occur around single stars
and members of binaries at the same rate, $\Lambda(r\p)$.

The only type of planet in this universe produces a transit signal of
amplitude $\delta_{\rm obs} = (r\p/R)^2$ when the host is a single
star.  Our naive observers try to measure the occurrence rate of these
planets.  They identify all the ``stars'' in their sample that appear
to be searchable for such a signal, given the size of their telescope.
These transits are detected within a certain maximum 
distance~\citep[see][]{pepper_using_2003,pepper_searching_2005}, which from
Equation~\ref{eq:S_N_thresh} scales as
\begin{equation}
  d_{\rm max} \propto \delta_{\rm obs} \cdot L_{\rm sys}^{1/2}.
  \label{eq:dmax}
\end{equation}
Assuming that stars are uniformly distributed in space, the number of
searchable stars $N_\star$ is then proportional to
\begin{equation}
    N_\star \propto n_i \delta_{\rm obs}^3 L_{\rm sys}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n_i$ is the number per unit volume of single stars, or
binary systems.

This selection process inadvertently admits some binaries.  Since the
binaries are twice as luminous as singles, they are included out to a
distance that is larger by a factor of $\sqrt{2}$, and if they have
transiting planets then the signals have an amplitude of
$(r\p/R)^2/2$.  To detect this half-strength signal, the observers
need the noise level to be lower by a factor of two, which means the
flux must be four times higher than that of a single star at distance
$d_0$.  This is true of single stars within a distance $d_0/2$, and
binaries within a distance $d_0\sqrt{2}/2$.
Furthermore the observers will consider any such shrunken signals to
represent planets of a different type, smaller in radius by a factor
of $\sqrt{2}$. These searchable volumes are illustrated in 
Figure~\ref{fig:model_1_volumes}.

Putting it all together, the {\it apparent} occurrence rate of planets
of radius $r\p$ will be
\begin{equation}
    \Lambda\a(r\p) = 
        \frac{\Lambda(r\p)~n\s d_0^3}
        {n\s d_0^3 + n\b (d_0\sqrt{2})^3}
\end{equation}
where $n\s$ and $n\b$ are the number densities of single stars and
binary stars, respectively, in the local neighborhood.  This apparent
rate is larger than the true rate by a factor
\begin{equation}
    \frac{\Lambda\a(r\p)}{\Lambda(r\p)} = 
        \frac{1}{1 + 2^{3/2}(n\b/n\s)}.
    \label{eq:correction_rp}
\end{equation}
The observers will also claim to have discovered a new type of planet
with radius $r\p/\sqrt{2}$ and an occurrence rate given by
\begin{equation}
\label{eq:correction_diluted_rp}
    \frac{\Lambda\a(r\p/\sqrt{2})}{\Lambda(r\p)}
    =
    \frac{2n\b (d_0\sqrt{2}/2)^3}
      {n\s (d_0/2)^3 + n\b (d_0\sqrt{2}/2)^3}
    =
    \frac{2\cdot2^{3/2}(n\b/n\s)}{1 + 2^{3/2}
    (n\b/n\s)}.
\end{equation}

@@

We can now assess the severity of these errors, given the
binary-to-single ratio $n\b/n\s$. For stars with masses from 0.7 to
1.3~$M_\odot$, \citet{raghavan_survey_2010} found the multiplicity
fraction to be 0.44, where multiplicity fraction is defined as
the fraction of systems in a volume-limited sample that are multiple.
Since we assume all multiple systems are binaries, this implies
\begin{equation}
    \frac{n\b}{n\s + n\b} \approx 0.44,
\end{equation}
which corresponds to a specific $n\b/n\s$.  Of course not all of these
binaries are close to being ``twin'' binaries as we have assumed in
our simple calculation. Perhaps only a tenth of them have pairs of
stars close enough in their basic properties to produce errors as
significant as those we have been considering.  Thus, an appropriate
estimate for $n\b/(n\s+n\b)$ is of order $0.05$, leading to
\begin{equation}
    \frac{\Lambda\a(r\p)}{\Lambda(r\p)} = 0.87,~~
    \frac{\Lambda\a(r\p/\sqrt{2})}{\Lambda(r\p)} = 0.26.
    \label{eq:ratios}
\end{equation}
All together the various effects produce biases of order 10\% in the
occurrence rates.  We will see that this level of error is
characteristic of many of our more complicated models, as well.

@@

There is a different way to conceptualize the preceding results that
will be useful in the subsequent discussion.  Here we broaden the
discussion to consider a spectrum of planet sizes, by introducing the
occurrence rate density
\begin{equation}
    \Gamma(r) \equiv \frac{d\Lambda}{dr},
\end{equation}
the average number of planets per star per unit planet radius.

To clarify our terminology, we define ``apparent properties'' as the
properties that one would infer about a system under the assumption
that the host star is single. For instance, the  ``apparent planet
radius'', $r\a$, is the planet radius that one would infer from a
transit signal, assuming that the host star is single.  The observed
transit depth is given by the ratio of the apparent planet radius and
the apparent stellar radius, $R\a$:
\begin{equation}
  \delta_{\rm obs}
  = \left(\frac{r\a}{R\a}\right)^2
  = \left[{r\over R_{\rm host}}\right]^2\times {L_{\rm host} \over
          L_{\rm sys}},
  \label{eq:delta_obs_general} 
\end{equation}
for $r$ the planet radius, $R_{\rm host}$ ($L_{\rm host}$) the host
star's radius (luminosity), and $L_{\rm sys}$ the entire system's
luminosity. This gives $r\a=r\p/\sqrt{2}$ in our ``twin binary''
model.

With this definition we may say that naive observers are measuring
$\Gamma\a(r\a)$, the apparent occurrence rate density of planets with
an apparent radius of $r\a$.  Analogous to
Equation~\ref{eq:occ_rate_simple},  
\begin{equation}
  \Gamma\a(r\a) = \frac{n_{\rm det}(r\a)}{N_\star(r\a)}
      \times \frac{1}{p_{\rm tra}},
\label{eq:apparent_rate_density}
\end{equation}
where $n_{\rm det}$ is the number of detected planets, per unit $r\a$,
with apparent radius $r\a$.  In
Equation~\ref{eq:apparent_rate_density}, we neglect the dependence on
stellar mass since all stars in this model have the same properties.
Since $\delta_{\rm obs}$ and $r\a$ are related by $R\a$, which is
simply assumed by an observer, measuring $\Gamma\a(r\a)$ is equivalent
to measuring the true occurrence rate density of {\it transit signals}
of amplitude $\delta_{\rm obs}$.

Let us evaluate the contributions to $n_{\rm det}$ from singles and
binaries.  As in the previous subsection, for each apparent planet
radius $r\a$, the observers select $N_\star(r\a)$ unresolved
point-sources on sky, based on Equation \ref{eq:dmax}.  If we let $\mu
\equiv N_\star^{\rm b}(r\a)/N_\star^0(r\a)$ be the ratio of the number
of searchable binary systems to single systems, then the number of
single and binary point-sources thought to be searchable are
\begin{equation}
  \frac{N_\star(r\a)}{1+\mu}~~{\rm and}~~\frac{\mu N_\star(r\a)}{1+\mu},
\end{equation}
respectively.  Earlier we showed that due to Malmquist bias, $\mu =
2^{3/2}(n\b/n\s)$.

We further stipulate that the true occurrence rate densities for
planets around single stars, primaries, and secondaries are
$\Gamma_0(r)$, $\Gamma_1(r)$, and $\Gamma_2(r)$, respectively.  
Then single stars contribute
\begin{equation}
  n_{\rm det}^0(r\a) = 
    {N_\star(r\a) \over {1+\mu}} p_{\rm tra} \Gamma_0(r\a)
  \label{eq:n0}
\end{equation}
planet detections per unit $r\a$.  We can similarly compute the number
of planet detections from binaries.  Since $r\p=\sqrt{2}r\a$ for
planets in twin binaries, 
\begin{equation}
  n_{\rm det}^1(r\a)\mathrm{d}r\a =
      {\mu N_\star(r\a) \over {1+\mu}} p_{\rm tra}
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1}
\end{equation}
detections come from primaries, and
\begin{equation}
  n_{\rm det}^2(r\a)\mathrm{d}r\a =
      {\mu N_\star(r\a) \over {1+\mu}} p_{\rm tra}
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n2}
\end{equation}
from secondaries.  Here the transit probability $p_{\rm tra}$ is the
same in all three cases because the stars are all identical.

>>>>

Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without knowing
which stars are binaries. Unresolved and unrecognized binary stars
give rise to systematic errors in planet occurrence rates, including
misclassified planets and miscounts in the number of searched stars.
The individual errors can have different signs, making it difficult to
anticipate the net effect on inferred occurrence rates. Here we use
simplified models of signal-to-noise limited transit surveys to try
and clarify the situation. We derive a formula for the apparent
occurrence rate density measured by an observer who falsely assumes
all stars are single. The formula depends on the binary fraction; the
mass function of the secondary stars; and the true occurrence of
planets around primaries, secondaries, and single stars. It also takes
into account the Malmquist bias by which binaries are over-represented
in flux-limited samples. Application of the formula to an idealized
{\it Kepler}-like survey shows that for planets larger than
$2r_\oplus$, the net systematic error is of order 10\%.  One
consequence is that unrecognized binaries cannot account for the
apparent discrepancy between hot Jupiter occurrence rates measured by
transit and RV surveys.  For smaller planets the errors are
potentially larger: the occurrence of Earth-sized planets could be
overestimated by as much as a factor of two. We also show that if
high-resolution imaging reveals a transit host star to be a binary,
the planet is usually more likely to orbit the primary star than the
secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

A lingering concern in occurrence rate studies is that in most cases,
investigators have assumed that all of the sources of light that were
monitored are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the point sources seen in a transit survey are
unresolved multiple-star systems, especially binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, in a binary there are two opportunities to detect
transiting planets, which could increase the overall number of
detections.

At the outset of this study it was not clear to us whether the neglect
of binaries was a serious issue at all, or even whether the net effect
of all the errors is positive or negative.  The goals of this study
were to provide a framework for dealing with these issues, and to
calculate a leading-order estimate for the net systematic effects.  In
this spirit, our models are idealized and analytic, and we do not
attempt a detailed correction of the results from {\it Kepler} or any
other real transit survey.

@@

Imagine a group of astronomers that wants to measure the mean number of
planets of a certain type per star of a certain type.  They perform a
wide-field photometric search for planets that transit, and discover
all for which
\begin{equation}
  \frac{\delta}{\sigma}
  >
  \left(\frac{{\rm S}}{{\rm N}}\right)_{\rm min}.
\label{eq:S_N_thresh}
\end{equation}
Here the signal S is the observed transit depth $\delta$, in units of
relative flux.  The noise N is the uncertainty in the determination of
the flux ratio of the star inside and outside of transits.  The
minimum signal-to-noise for detection, $({\rm S}/{\rm N})_{\rm min}$,
depends on the properties of the telescope.  We will assume throughout
this paper that the photometric uncertainty scales as the inverse
square root of the number of photons collected from the source, {\it
i.e.}, the dependence on stellar properties is $\sigma \propto (L_{\rm
tot}/d^2)^{-1/2}$, where $L_{\rm tot}$ is the total luminosity of the
stellar system and $d$ is its distance from Earth.  The observed
signal amplitude $\delta$ depends on the fraction of the total
luminosity that comes from the planet-hosting star, since light from
unresolved stellar companions is not affected by the planet's transit.
% The transit duration also affects detectability, but we omit it in
% this work for brevity.

The astronomers analyze their data assuming that all stars are
single.  In particular they do not have accurate enough parallaxes to
tell that some of the stars are apparently overluminous.  They want to
compute the number of planets with size $r$ per star of mass $M$.  If
the total number of stars for which such planets could be detected
($N_\star$) and the number of detected planets ($N_{\rm det}$) are
both large, then the astronomers' estimate for the occurrence rate
($\Lambda$) is simply
\begin{equation}
  \Lambda = \frac{N_{\rm det}}{N_\star}
                    \times \frac{1}{p_{\rm tra}},
\label{eq:occ_rate_simple}
\end{equation}
where the geometric transit probability, $p_{\rm tra}$, accounts for
the fact that most planetary orbits are not aligned close enough with
our line of sight to produce transits.

@@

We will focus exclusively on problems that arise from the fact that
many stars exist in gravitationally bound binary systems.  Even with
this narrow focus, there are numerous sources of error.  All three of
the quantities in Equation~\ref{eq:occ_rate_simple} are subject to
observational bias:
\begin{enumerate}
%    
    \item The detected planet count, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size will be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The apparent number of stars that were searched, $N_\star$,
      is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were searched for transiting
          planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be suitably small and bright to detect the
          desired type of planet are in fact binaries for which the
          amplitude of the photometric signal would be reduced to an
          undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$, which scales with the
      stellar mean density as $\rho^{-1/3}$, is biased because the
      planet-hosting star could be misclassified.
%    
\end{enumerate}

Two other problems that may arise are not represented in
Equation~\ref{eq:occ_rate_simple}.  One is astrophysical: the true
occurrence rate of the desired type of planet may depend on whether
the host is a single star, the primary star of a binary, or the
secondary star of a binary.  Such differences could be caused by the
requirement for long-term dynamical stability, or differences in the
planet formation process.  When the search sample includes both
singles and binaries, the detected planets are thereby drawn from
different occurrence 
distributions~\citep[see][]{
    wang_occurrence_2015,kraus_impact_2016}.

The other problem is an observational effect. In the searched sample,
the ratio between the number of binary and single stars will differ
from that in a volume-limited sample due to a type of Malmquist bias.
The total luminosity of a binary is larger than the luminosity of
either the primary or secondary star.  This means that for transit
signals of a given amplitude, the binaries are searchable at greater
distances from the Earth.  Thus for any planet and star pair, adding a
stellar companion both helps and hinders transit detection: the extra
photons help by lowering the noise, and hinder by diminishing the
observed transit depth.
% Accounting for each effect requires some care.

@@

Given all of the confusing and opposing sources of error, we will
proceed in stages. We start from a barebones model in which everything
can be written down on the back of a napkin, and build up to an
analytic model allowing for generality in the distribution of the
binaries and the planets they host.

@@

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{
        Searchable volumes for single stars ({\it left}) and twin
        binaries ({\it right}), assuming that all stars have the same
        mass, radius, and luminosity, and that all planets have the
        same radius $r\p$.
        {\it Top:} At an apparent radius $r\a=r\p$, the
        binarity-ignoring observers search twin binaries out to a
        distance $\sqrt{2}$ times greater than they search single
        stars.  Because the transit depths are diluted, there are no 
        planets in twin binaries with $r\a=r\p$.
        {\it Bottom:} At $r\a=r\p/\sqrt{2}$, the observed transit
        depths are smaller by a factor of two, so the maximum
        searchable distances are half those at $r\a=r\p$.  The only
        detected planets with $r\a=r\p/\sqrt{2}$ orbit twin binaries.
    }
    \label{fig:model_1_volumes}
\end{figure}

Since the effects of binarity are most pronounced when the two stellar
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Some fraction of them exist in twin binaries. All planets have
size $r\p$.  In this scenario, stellar radii are never misclassified
because the combined light of a binary has the same color and spectrum
as a single star.  We further assume that planets occur around single
stars and members of binaries at the same rate, $\Lambda(r\p)$.

Our naive observers detect transiting planets with two different
depths. Since 
\begin{equation}
    \delta = \left(\frac{r\p}{R}\right)^2 \frac{L}{L_{\rm tot}},
    \label{eq:delta_obs_general} 
\end{equation}
one class of transit signal, with depth $(r\p/R)^2$, occurs in single
star systems. The other class, with half the depth, occurs in twin
binary systems, leading to apparent planet radii $\sqrt{2}$ smaller
than the true radii.

To calculate the occurrence rate of each planet type, our observers
identify the ``stars'' from their survey that appear to have been
searched for signals of each depth.  These transits are detected
within a certain maximum
distance~\citep[see][]{pepper_using_2003,pepper_searching_2005}, which
from Equation~\ref{eq:S_N_thresh} scales as
\begin{equation}
  d_{\rm max} \propto \delta \cdot L_{\rm tot}^{1/2}.
  \label{eq:dmax}
\end{equation}
Assuming that stars are uniformly distributed in space, the number of
searchable stars $N_\star$ is proportional to
\begin{equation}
    N_\star \propto n \delta^3 L_{\rm tot}^{3/2},
\label{eq:N_searchable_prop}
\end{equation}
where $n$ is the number per unit volume of single or binary star
systems.

When selecting sources that appear to have been searched, the
observers admit some binaries.  Since the binaries are twice as
luminous as singles, at fixed $\delta$ they are included out to a
distance that is larger by a factor of $\sqrt{2}$.  To detect the
half-strength signal, the observers need the noise level to be lower
by a factor of two, which means the flux must be four times higher
than that of a single star at distance $d_0$.  This is true of single
stars within a distance $d_0/2$, and binaries within a distance
$d_0\sqrt{2}/2$.  The searchable volumes are illustrated in
Figure~\ref{fig:model_1_volumes}.

Putting it all together, the {\it apparent} occurrence rate
$\Lambda\a$ of planets of radius $r\p$ will be
\begin{equation}
    \Lambda\a(r\p) = 
        \frac{\Lambda(r\p) \, n\s d_0^3}
        {n\s d_0^3 + n\b (d_0\sqrt{2})^3}
    \label{eq:rate_at_rp}
\end{equation}
where $n\s$ and $n\b$ are the number densities of single and binary
star systems, respectively.  This apparent rate is larger than the
true rate by a factor
\begin{equation}
    \frac{\Lambda\a(r\p)}{\Lambda(r\p)} = 
        \frac{1}{1 + 2^{3/2}(n\b/n\s)}.
    \label{eq:correction_rp}
\end{equation}
The observers will also claim to have discovered a class of planet
with radius $r\p/\sqrt{2}$ and an occurrence rate given by
\begin{equation}
    \frac{\Lambda\a(r\p/\sqrt{2})}{\Lambda(r\p)}
    =
    \frac{2n\b (d_0\sqrt{2}/2)^3}
      {n\s (d_0/2)^3 + n\b (d_0\sqrt{2}/2)^3}
    =
    \frac{2\cdot2^{3/2}(n\b/n\s)}{1 + 2^{3/2}
    (n\b/n\s)}.
    \label{eq:correction_diluted_rp}
\end{equation}

@@

We can now assess the severity of these errors, given the
binary-to-single ratio $n\b/n\s$. For stars with masses from 0.7 to
1.3~$M_\odot$, \citet{raghavan_survey_2010} found the multiplicity
fraction~--~the fraction of systems in a volume-limited sample that
are multiple~--~to be 0.44.  Assuming all multiple systems are
binaries, this means
\begin{equation}
    \frac{n\b}{n\s + n\b} \approx 0.44.
\end{equation}
Of course not all of these binaries are ``twin'' binaries as we have
assumed in our simple calculation. Perhaps only a tenth of the
binaries have pairs of stars close enough in their basic properties to
produce errors as significant as those we have been considering. In
detail, this fraction depends on the binary mass ratio distribution
and stellar mass-luminosity relation (see
Section~\ref{sec:general_formula}).  Taking a leading-order estimate
of $n\b/(n\s+n\b)\approx0.05$, we find
\begin{equation}
    \frac{\Lambda\a(r\p)}{\Lambda(r\p)} = 0.87,~~
    \frac{\Lambda\a(r\p/\sqrt{2})}{\Lambda(r\p)} = 0.26.
    \label{eq:ratios}
\end{equation}
All together, the various effects produce biases of order 10\% in the
occurrence rates.  We will see that this level of error is
characteristic of many of our more complicated models as well.

@@

There is a different way to conceptualize the preceding results that
will be useful in the upcoming discussion. We now consider a spectrum
of planet sizes, $r$, and introduce the occurrence rate density
\begin{equation}
    \Gamma(r) \equiv \frac{{\rm d}\Lambda}{{\rm d}r},
    \label{eq:rate_density_defn}
\end{equation}
the number of planets per star per unit planet radius.

The naive observers are in fact measuring $\Gamma\a(r\a)$, the {\it
apparent} occurrence rate density~--~the number of planets per star
per unit apparent radius $r\a$.  The ``apparent planet radius'' is the
radius inferred from the transit depth under the assumption that the
host star is single. The ``apparent stellar radius'' $R\a$ is
analogous.  The observed transit depth can be written $\delta =
(r\a/R\a)^2$, which from Equation~\ref{eq:delta_obs_general} gives
$r\a=r\p/\sqrt{2}$ for planets in twin binaries.  $\Gamma\a(r\a)$ is
the same as the {\it true} rate density of transit signals of
amplitude $\delta$.

We will now calculate the apparent rate density by evaluating the star
and planet counts for signals coming from single stars, primary stars
in binaries, and secondary stars in binaries.  Let $\mu$ be the ratio
between the number of searchable binary systems and the number of
searchable single systems at a particular transit depth. The number of
single and binary sources in the searched sample are then
\begin{equation}
  \frac{N_\star(r\a)}{1+\mu}~~{\rm and}~~\frac{\mu N_\star(r\a)}{1+\mu},
  \label{eq:N_searched}
\end{equation}
respectively.  Earlier we showed that due to Malmquist bias, $\mu =
2^{3/2}(n\b/n\s)$.

Suppose the occurrence rate densities for planets around
single stars, primaries, and secondaries are $\Gamma_0(r),
\Gamma_1(r)$, and $\Gamma_2(r)$, respectively. For a single planet
size,
\begin{equation}
    \Gamma_i(r) = \Lambda(r\p) \cdot \hat{\delta}(r-r\p),~~{\rm for}
    \ i\in\{0,1,2\},
    \label{eq:one_planet}
\end{equation} 
where $\hat{\delta}$ is the Dirac delta function.
The number of planets detected from singles, primaries, and
secondaries are then
\begin{align}
  n_{\rm det}^0(r\a)\mathrm{d}r\a &= 
      \frac{N_\star(r\a)}{1+\mu} p_{\rm tra}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  n_{\rm det}^1(r\a)\mathrm{d}r\a &= 
      \frac{\mu N_\star(r\a)}{1+\mu} p_{\rm tra}
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  n_{\rm det}^2(r\a)\mathrm{d}r\a &=
      \frac{\mu N_\star(r\a)}{1+\mu} p_{\rm tra}
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n2}
\end{align}
where $n_{\rm det}^i$ is the number of detections per unit apparent
radius for each respective type of star.

Dividing the number of detections by the number of sources,
the apparent rate density reported when ignoring binarity is
\begin{equation}
  \Gamma\a(r\a) =
    \frac{1}{1+\mu}\,\Gamma_0(r\a) +
    \frac{\mu}{1+\mu}\,\sqrt{2}\,\Gamma_1(\sqrt{2}r\a) +
    \frac{\mu}{1+\mu}\,\sqrt{2}\,\Gamma_2(\sqrt{2}r\a).
	\label{eq:formula_twin_binary}
\end{equation}
If we insert Equation~\ref{eq:one_planet} and integrate over
apparent radius, this reproduces Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}.

\begin{comment}
  and inputting Equation~\ref{eq:one_planet}, we find
  \begin{align}
      \Gamma\a(r\a) &= \Lambda(r\p) \left[
      \frac{1}{1+\mu} \cdot
      \hat{\delta}(r\a-r\p)  +
      \frac{2\mu}{1+\mu} \cdot
      \delta\left(r\a-\frac{r\p}{\sqrt{2}} \right) \right].
      \label{eq:model_1_apparent_rate_density}
  \end{align}
  Integrating over $r\a$, this yields
  \begin{equation}
    \Lambda\a(r_{\rm p})=\frac{Z}{1+\mu},\quad
    \Lambda\a(r_{\rm p}/\sqrt{2})=\frac{2\mu Z}{1+\mu},
  \end{equation}
  which reproduce Equations~\ref{eq:correction_rp}
  and~\ref{eq:correction_diluted_rp}.
\end{comment}

\begin{comment}
    The ratio of the apparent to single star rate
    densities,
    \begin{equation}
        X_\Gamma(r) \equiv \left. \frac{\Gamma\a(r\a)}{\Gamma_0(r)}
          \right|_{r\a = r},
    \end{equation}
    can then be evaluated as a ``correction factor''.  At the true and
    diluted planet radii, $X_\Gamma(r\p)=1/(1+\mu)$ and
    $X_\Gamma(r\p/\sqrt{2})=2\mu/(1+\mu)$.  This yields a correction
    $X_\Gamma$ of 0.87~(0.76) if the binary fraction is 0.05~(0.10).
    All together the various effects produce biases of order 10\% in the
    occurrence rates.  
\end{comment}
